---
title: "smart watch price prediction"
author: "Eva Ssozi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This project aims to develop a model to predict the price of smartwatches. The dataset was obtained from Kaggle at: https://www.kaggle.com/datasets/rkiattisak/smart-watch-prices

```{r}
#libraries required:
library(tidyverse)
library(readxl)
library(readr)
library(esquisse)
library(caret)

```

# Part 1: Tidying up data and EDA
```{r}
#read in the dataset: there are rows with N/A, Not specified
df <- read.csv("Smart watch prices.csv", header = T, sep = ",", strip.white = T, na.strings=c(" ","","NA","N/A","Not specified")) #removing leading and trailing white spaces, and reading NAs as NAs

#inspect structure of imported data:
dim(df) #379 rows, 13 columns
str(df)
```



-Use table function to take a closer look at each column contents, so that we know how to approach cleaning:
```{r}
apply(df, 2, table)
```

From the previous table() inspection, there are rows with unlimited (for battery life).

Upon further inspection, battery life column has only 3 rows with character type of "unlimited." We can create bins (ranges) from this column so that we can model it as a categorical variable. This column is in days,  the difference between 1 day and 2 days in 24hours, so we cannot make the bins too big, since this has an effect on how the watch is used and who buys it. Min is 1 day, Max is unlimited, so lets create bins of 3days, that is, 1-3, 4-6, 7-9, etc.

```{r}
#inspect battery life column:
battlife <- as.data.frame(df$Battery.Life..days.)
table(battlife)
```

```{r}
#battery life column needs to be numerical, so replace 48 hours with 2 days and Unlimited with arbitrary placeholder number which is just greater than max.

battlife$timelength <- battlife$`df$Battery.Life..days.`#first copy column

#48 hours is 2 days:
battlife$timelength <- ifelse(battlife$timelength == "48 hours", "2", battlife$timelength)

#Unlimited is placeholder 1000:
battlife$timelength <- ifelse(battlife$timelength == "Unlimited", "1000", battlife$timelength)

#inspect result:
table(battlife$timelength)

```

```{r}
#creating bins:
battlife$numbers <- as.numeric(battlife$timelength)

battlife <- battlife %>% mutate(new_bin = cut(numbers, breaks=c(0,3,6,9,12,15,18,21,24,27,30,33,36,39,42,45,48,51,54,57,60,63,66,69,72,75,78,81,84,90)))

#export for excel edit:
write.csv(battlife, file="battery_life.csv", row.names = F)
```


```{r}
#read in excel sheet with ranges for battery life added:
battery_ranges <- read_excel("battery_life_ranges.xlsx", sheet = "battery_life")
battery_ranges$numbers <- as.numeric(battery_ranges$numbers)
battery_ranges <- battery_ranges %>% select(numbers, Range) %>% distinct() #29
```

```{r}
#add range to battlife:
battlife$timelength <- as.numeric(battlife$timelength)

battlife <- battlife %>%
  left_join(battery_ranges, by = c("numbers"="numbers")) #379 rows, so okay
```

```{r}
#add battery life range to df:
df$battery_range_days <- battlife$Range #379, so okay.
#check:
check <- df %>% select(Battery.Life..days., battery_range_days) %>% distinct()
```

```{r}
#remove "$" and "," in price column:
df$cleanPrice_usd <- gsub("\\$", "", df$Price..USD.)
df$cleanPrice_usd <- gsub(",", "", df$cleanPrice_usd)
#to numeric:
df$cleanPrice_usd <- as.numeric(df$cleanPrice_usd)
```

```{r}
#filter out rows with blanks:
df2 <- na.omit(df) #373 rows left out of 379
```

```{r}
#remove old battery life and price related columns:
df2 <- df2 %>% select(-c("Battery.Life..days.","Price..USD."))
```

-columns need to be in right format, that is, numerical and factor/categorical
```{r}
#categorical columns:
df2 <- df2 %>%
  mutate_if(is.character, as.factor)
```

```{r}
#check models since it has 134 levels:
checkmodels <- df2 %>%
  select(Brand, Model, cleanPrice_usd) %>%
  distinct() %>% arrange(Brand)
```

there are some combinations of brands, models and prices that don't seem to be right, for example the models under the Zepp brand but we can't exclude them based on guesswork, so left as is.

# Using simple graphics to visualize the variables before proceeding to predict prices
```{r}
brandgg <- df2 %>%
  ggplot(aes(x = fct_infreq(Brand))) +
  geom_bar(fill = "blue") +
  theme(legend.position = "bottom") +
  labs(x = "Brand", y = "Count") +
  theme(axis.text.x = element_text(angle = 90))

brandgg
```


```{r}
osgg <- df2 %>%
  ggplot(aes(x = fct_infreq(Operating.System))) +
  geom_bar(fill = "blue") +
  theme(legend.position = "bottom") +
  labs(x = "Operating System", y = "Count") +
  theme(axis.text.x = element_text(angle = 90))

osgg
```


```{r}
conectgg <- df2 %>%
  ggplot(aes(x = fct_infreq(Connectivity))) +
  geom_bar(fill = "blue") +
  theme(legend.position = "bottom") +
  labs(x = "Connectivity", y = "Count") +
  theme(axis.text.x = element_text(angle = 90))

conectgg
```


```{r}
disptygg <- df2 %>%
  ggplot(aes(x = fct_infreq(Display.Type))) +
  geom_bar(fill = "blue") +
  theme(legend.position = "bottom") +
  labs(x = "Display Type", y = "Count") +
  theme(axis.text.x = element_text(angle = 90))

disptygg
```


```{r}
resolgg <- df2 %>%
  ggplot(aes(x = fct_infreq(Resolution))) +
  geom_bar(fill = "blue") +
  theme(legend.position = "bottom") +
  labs(x = "Resolution", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0))

resolgg
```

```{r}
heartmongg <- df2 %>%
  ggplot(aes(x = fct_infreq(Heart.Rate.Monitor))) +
  geom_bar(fill = "blue") +
  theme(legend.position = "bottom") +
  labs(x = "Heart Rate Monitor", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0))

heartmongg
```

```{r}
gpsgg <- df2 %>%
  ggplot(aes(x = fct_infreq(GPS))) +
  geom_bar(fill = "blue") +
  theme(legend.position = "bottom") +
  labs(x = "GPS", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0))

gpsgg
```

```{r}
battrangegg <- df2 %>%
  ggplot(aes(x = fct_infreq(battery_range_days))) +
  geom_bar(fill = "blue") +
  theme(legend.position = "bottom") +
  labs(x = "Battery range", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0))

battrangegg
```

```{r}
#simple statistics for all variables:
summary(df2)
```
# **Price prediction**

## **Mulitiple linear regression model:**

-We see that the heart rate monitor column has only one variable, "Yes", so this feature can be excluded for the price prediction steps.
```{r}
df3 <- df2 %>% select(-Heart.Rate.Monitor) # remove heart rate monitor colum
#structure of df3:
dim(df3)
str(df3)
```

-Investigate multicollinearity among the different 12 features/predictor variables (minus the CleanPrice variable which is the target variable):

* Multicollinearity:  
A pairwise scatter plot is plotted to identify any multicollinearity in the data. This is an important step because the regression analysis process assumes that the predictor variables are independent of each other. The pairwise plot enables us to observe any pair of predictor variables that might be highly correlated, and require further investigation to decide which variable should be included/elminated from the analysis.  

```{r}
#Standardization of quantitative variables, excluding the factor variables:
df3_st <- df3 %>%
  mutate_at(c("Display.Size..inches.","Water.Resistance..meters."),~(scale(.) %>%
                          as.vector))
```

```{r}
pairs(df3_st)
```

```{r}
#to view correlation coefficients:
numeric_df3_st <- df3_st %>% select(Display.Size..inches., Water.Resistance..meters., cleanPrice_usd)
cor(numeric_df3_st) # there is no high collinearity between the numeric variables
```

There is no mulitcollinearity between the numeric variables, so we can proceed with regression analysis:


```{r}
#splitting dataset into train and testing sets
set.seed(1)
index <- createDataPartition(df3_st$Operating.System, times=1, p=0.7, list=F)
train <- df3_st[index,]
test <- df3_st[-index,]
```

-The categorical variables have many levels compared to the size of the dataset, so we explore building a model without use of training and testing sets.
```{r}
lm_1 <- lm(cleanPrice_usd ~., data = df3_st)
summary(lm_1) #multiple r squared: 0.9914, adjusted R-squared: 0.9807
```

-Testing lm_1 using entire data set:
```{r}
lm_1.pred <- predict(lm_1, newdata = df3_st)

#calculating mse for test set:
mean((lm_1.pred-(df3_st$cleanPrice_usd))^2)
```


## Lasso regularisation model: 

The Lasso method employs an inbuilt constraint instead of calculating penalties (for example cp and bic) separately.
A  model is fitted, with 100 lambda values ranging from 10^10 to 10^-2. The goal is to select the lambda using cross validation, as a way to minimize the test error. The mean square error is calculated from predicted values using the selected lambda.

```{r}
x<- model.matrix(cleanPrice_usd~.,df3_st)[,-12] #creates matrix of predictor variables for all observations
y<- df3_st$cleanPrice_usd #column vector of response variables
```

```{r}
library(glmnet)
grid <- 10^seq(10,-2, length = 100)
lassomodel <- glmnet(x, y, alpha=1, lambda=grid) #alpha = 1 is for lasso
plot(lassomodel)
```

The figure below illustrates the variation of the mean squared error (as calculated from the train data) with log(lambda). We choose our best lambda as the lambda associated with the minimum mean square error in the train set. This lambda is then used to predict response variables with the test set data.

```{r}
set.seed(1)
crossvalid.out <- cv.glmnet(x, y, alpha=1)
plot(crossvalid.out)
bestlambda <- crossvalid.out$lambda.min #selecting the best lambda, that is at minimum error, 1.9358
#calculating predicted values using the data:
lassopred <- predict(lassomodel, s=bestlambda, newx = x)
```


```{r}
out <- glmnet(x,y,alpha =1,lambda = grid)
lasso.coef <- predict(out, type = "coefficients", s=bestlambda)[1:141,]
lasso.coef
lasso.coef[lasso.coef!=0]
```

```{r}
#calculating mean square error from lasso regularisation:
mean((lassopred-y)^2)
```

- The mulitple linear regression model without regularisation gave a lower mse.

